---
title: "Diameter_analysis_Ultem(Stratasys)"
author: "Rohan"
date: "2026-07-09"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

## R Markdown

## Filament Diameter Analysis — Ultem (Stratasys)

This document analyzes the diameter variation of a ~51m spool of Ultem filament 
(Stratasys) measured using a bench test rig..

Two orthogonal Hall-effect diameter sensors (TMAG5273) and a Lika A41H incremental 
encoder were used to measure filament diameter and length simultaneously. 
Sensor 1 (0x78) and Sensor 2 (0x22) are positioned 60mm and 120mm downstream 
of the encoder respectively.

The analysis examines diameter variation between the two sensors, ovality along 
the spool length, and deviation from the nominal filament diameter of 1.75mm.

```{r import Data}
files<-list.files(path= "../Logged_data",
                  pattern = "*.csv",
                  full.names = TRUE)

data<- files %>% 
  map(read.csv) %>% 
  bind_rows()#combining all the rows



data_r<- data %>% 
  distinct(length_mm,.keep_all=TRUE) %>%  # removing duplicates or cases where the spool was stationary
  mutate(
    pos_sensor_1= length_mm-60, # distance between point of measurements and logged measurement for sensor 1
    pos_sensor_2=length_mm-120,# distance between point of measurements and logged measurement for sensor 2
    ovality= abs(diameter1_mm-diameter2_mm),
    length_m=length_mm/1000,
                 segment=cut(length_m,breaks=seq(0,52,by=5.2),include.lowest=TRUE))
    
```

## Plot 1 — Diameter variation between sensors
```{r box_plot_diameter_variation}
ggplot(data_r)+
  geom_boxplot(aes(x="Sensor_1", y=diameter1_mm,fill= "Sensor_1"))+
  geom_boxplot(aes(x="Sensor_2",y=diameter2_mm,fill = "Sensor_2"))+
  labs(title = "Diameter Variation", x="Sensor",y="Diameter(mm)",fill= "Diameter_Sensor")
```
## Plot 2 — Diameter variation between sensors (Density plot)
```{r plot_diameter_density}
data_density <- data_r %>%
  pivot_longer(
    cols = c(diameter1_mm, diameter2_mm),
    names_to = "sensor",
    values_to = "diameter"
  ) %>%
  mutate(sensor = recode(sensor,
                         "diameter1_mm" = "Sensor 1",
                         "diameter2_mm" = "Sensor 2"))

ggplot(data_density, aes(x = diameter, fill = sensor)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 1.75, linetype = "dashed", colour = "red") +
  labs(title = "Diameter Distribution by Sensor",
       x = "Filament Diameter (mm)",
       y = "Statistical Density") +
  xlim(1.65, 1.85)
```

## Plot 3 — Ovality along the length of the filament measured
```{r plot_ovality}
ggplot(data_r, aes(x=segment,y=ovality))+
  geom_boxplot(fill="steelblue")+
  geom_smooth()+
  labs(title = "Ovality along the length",x="Length(m)",y="Ovality(mm)")+
theme(axis.text.x = element_text(angle=45,hjust=1))
```

## Plot 4 — Deviation from 1.75mm
```{r plot_deviation_from_1.75}

data_long<- data_r %>% 
  mutate(
    dev_sensor1=diameter1_mm-1.75,
    dev_sensor2=diameter2_mm-1.75) %>% 
  pivot_longer(cols=c(dev_sensor1,dev_sensor2),
               names_to="sensor",
               values_to = "deviation") %>% 
  mutate(sensor=recode(sensor,
                       "dev_sensor1"= "Sensor_1",
                       "dev_sensor2"="Sensor_2"))


ggplot(data_long,aes(x=length_m,y=deviation,colour = sensor))+
  geom_point(alpha=0.2)+
  geom_smooth()+
  geom_hline(yintercept = 0,linetype="dashed",colour="red")+
  labs(title="Deviation from the diameter of 1.75mm along length of filament",x="Length(m)", y="Deviation(mm)",color= "Deviation from sensors")
  
```
